### 2016-07-07 image_setup.tex
Starting with overlay fs

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen} for better handling of
long-runnning processes. When writing the image either to a memory card
or storing it remotely \todot{Are we going to store the image
into a memory card or remotely in a URL? Since this is not completely clear
yet.}, all these change that has been made to the image so far will exists in
all \ac{Raspi}s.

>>>>

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen} for better handling of
long-runnning processes. When writing the image to a memory card, all the
changes that have been made to the image so far, will exist in all
\ac{Raspi}s after fetching it.

### 2016-07-07 overlay_fs.tex
Starting with overlay fs

\label{sec:ovelay_fs}

\subsection{Installation an overlaying filesystem}

Our goal was to enable the users of the testbed to change however the
\ac{Raspi} are configured while still being able to revert those changes easily.

We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.

A way to do this is using stacked filesystem. Stacking filesystem enables
a filesystem to overlay another filesystem. Whenever a file is accessed,
the upper filesystem will forward the request to the lower filesystem in
case it does not have it itself. If it has it, it will simpy return the file.

This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
be changable. Some programs needs to change files and the testbed admin might
want to change configurations. To enable this, it is possible to
overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
the regular root. Thus, reading a file may return a file from the lower filesytem,
but if it is stored, it will be saved in the upper filesystem. Accessing this file
again will return the stored file from the upper layer.

%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
%mount their root filesystem as 
%
%We do this using an overlay filesystem (overlayfs).
%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
%
%
%Overlayfs enables filesystems
%to be stacked on top of each other. We will be using this to make a lower
%filesystem and an upper filesystem. The lower filesystem will be our
%traditional filesystem and the upper filesystem will be mounted in tmpfs
%(i.e. mounted in \ac{RAM}).
%
%Thus, when a file is requsted from stacked filesystem, it will first attempt
%to fetch the file in the upper filesystem if it does not exists there, it will
%request it from the lower filessytem.

Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
can setup the overlay filesystem (overlayfs).

There already exists implementations overlaying the root filesystem. We will
use an existing implementation
available at github: \cite{overlayroot}.

We installed git in a previous step, so lets clone the repository. The below
code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
will disappear when the machine is rebooted.

>>>>

\label{sec:ovelay_fs}
In principle, our procedure modifies the image file only once in
the testbed administrator \ac{PC} when its setup is made. Also, keeping
this image in the \ac{Raspi}s provides the same initial system for all
the devices. This system is later distributed locally to a memory card
in each device. However, if we do not make any further modifications during
the image setup, any files created after the initial boot of a \ac{Raspi}
will remain in the memory card. This is cumbersome to maintain since
the size of the memory card is relatively reduced (8 GB) and there
might be various users utilizing the testbed. Also, different testbed
users could be interested in running their experiments in a fresh
rebooted system with the original customized image. Nevertheless, we
emphasize that this step is not necessary if the tesbed administrator
wants to consider only persistent storage for its devices. A use case
for this scenario could be a single user for the testbed or when a testbed
administrator only wants to setup a few \ac{Raspi}s.

Therefor, if both persistent and non-persistent storage are required for
the \ac{Raspi}s, here we present the steps to setup an overlay filesystem.
This type of filesystem enables an \textit{upper} filesystem to overlay
into a \textit{lower} filesystem. Whenever a file is requested, the upper
filesystem will forward the request to the lower filesystem in case it
does not have it itself. If the upper filesystem has the requested file,
it will simpy return the file. This idea can be used in our setup to mount
the root filesystem (i.e. Raspbian lite) in the \ac{Raspi}s during startup
as read-only filesystem. Thus, on the one hand, the image configuration files
will remain after a reboot and will not be modifiable. On the other hand,
to enable the possibility of non-persistent changes, we overlay the upper
filesystem that is mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp}
as read/writable on top of the lower root filesystem. Thus, reading a file
may return a file from the lower filesytem, but if it is stored, it will
be saved in the upper filesystem. Accessing this file again will return
the stored file from the upper layer. After a reboot, all the stored files
in the upper filesystem will be erasured.

\subsection{Overlaying Filesystem Installation}

Assuming that we are still using the \texttt{chroot} commands into the
Raspbian lite root filesystem for installing packages, we can setup the
overlay filesystem at this point of the procedure. There already exists
implementations overlaying the root filesystem. We use an implementation
available at the Git repository in\cite{overlayroot}. Since we have
installed git in a previous step, we clone the repository. The
command block below stores it in \texttt{/tmp} which is really mounted in
\ac{RAM}. Then, all the files stored here will disappear a \ac{Raspi} is
rebooted.
